Mixed language processing increases cross-language phonetic transfer in Bengali-English bilinguals

### 2022-06-01 draft_BLC/blc_reviewer_response.txt

AR: Thank you for pointing this out. The statistical analyses in this study is within-participant, comparing participants' IE production in a unilingual condition to their own mixed productions. The baseline Bengali data was used as a heuristic tool to understand how the direction of shift relates to Bengali vowel categories. We had initially used data from Dutta (2021), which was a subset of speech data from the SHRUTI corpus[i], extracted for another study. However, we agree with the reviewer's assessment that the small dataset an lack of gender balance makes it not ideal representative of Bengali vowel production. We have now discarded the Dutta (2021) data, and instead used the entirety of the SHRUTI corpus to get average vowel positions. This corpus contains speech from 34 speakers of Bengali (includes both male and female speakers) and is the largest available corpus of Western Bengali speech that we are aware of. We have elaborated on details of the corpus and speakers on page 16. While the relative vowel positions we observe have not changed as a result of using this larger dataset, we believe that this dataset is more 
representative of average Bengali vowels, and address the reviewer's concern that the observed results might be an artifact of the particular speech data used earlier.

[i] Das, B., Mandal, S., & Mitra, P. (2011, October). Bengali speech corpus for continuous auutomatic speech recognition system. In 2011 International conference on speech database and assessments (Oriental COCOSDA) (pp. 51-55). IEEE.

2. Participants.
a) How was the sample size calculated?
AR: This was based on resource constraints and participant availability based on our inclusion criteria (c.f. response to comment 2 of reviewer 1). We have elaborated on this process and reported our inclusion criteria in page 20 of the revised manuscript.

b) Participants should be described better, especially for what concerns their L2 experience. Currently, there is no information about their proficiency level, as well as their degree of exposure and use of their two languages.
AR: Thanks for this suggestion. We have added the suggested content to the manuscript on page 19; c.f. response to comment 2 of reviewer 1.

3. Materials, how many of the 20 words contained each of the two target vowels? Please specify this information in the text.
AR: We have added this information on page  23:
"Twenty monosyllabic English words: ten containing the vowel ʌ and ten containing æ, were selected as target words."

4. Procedures.

a) If I correctly understand, each word was repeated 4 times in each task (i.e., the same word was produced for a total of 8 times by each participant). If this is the case, can the authors exclude any repetition effect? And why did they opt for a so massive repetition, instead of, e.g., increasing the number of items?

AR: Our choice of items was constrained by phonological considerations, to ensure that coarticulatory effects that are known to affect vowel acoustics, such as identity of the pre- and post-vocalic consonant, were minimal, and controlled across items. Thus, we only used words where the pre-vocalic consonant was a voiced plosive /b/ or /d/, and had a balanced set of post-vocalic consonants across the set of items. Taking into consideration lexical frequency (we only included common words to avoid hesitation while reading) and loanword status (we excluded lexicalized loanwords, e.g. "cup") allowed for 10 target items. We opted for a larger number of repetitions because (i) we wanted multiple responses per participant per item to control for random variability and potential production errors, and (ii) to ensure enough data points for robust statistical analyses. Given that items were presented in random order, and participants were presented with intermediate distracter tasks during the cued picture-naming which relies on language priming, we do not think that repetition effects would come into play in this design.

b) Cued-picture naming: why was the picture presented for a so short time (50 ms)? What words were used as language cue and as target?

AR: In this task the picture, named in either Bengali or English, serves to prime the participant for that particular language. The following target word then either complies with the prime (unilingual condition) or deviates from it (inducing switching; mixed condition). Since participants are already in a bilingual language mode, the effects of such priming, if any, are expected to be highly localized (c.f. discussion in section 1.2 and 1.3, page 9). We presented the picture for a short time to avoid a pause between the prime and the target that would potentially undo the effect of priming, or introduce unwarranted planning effects. Another consideration was for productions in both tasks to mimic the time course and prosody of naturalistic speech as much as possible. Thus, we wanted to avoid long pauses or unnatural intonation that might arise from producing the target as an isolated word.

We have specified the words used as langauge cue, and provided a link to the list of target words, in the text (page 25).

5. Analyses.
a) To analyze the changes at different times during the vowel production, I would suggest to run a single model in which the time of measurement (5%, 15%, 25%, 35%, 45%) is entered as a numeric predictor (from 1 to 5). This will reduce the number of statistical models to run and, more importantly, will allow to directly study the time course of the changes.
AR: This is an excellent suggestion, thank you! We have re-run the analysis as suggested.

b) p. 22 ls 21-23. The authors should specify which interactions they are referring to and report the results of the statistical tests.
AR: While we have now re-written this section to reflect the revised analysis, we have reported the effect sizes, standard errors, and p-values associated with all interaction terms of interest, and reported chi-square, degrees of freedom, and p-values for all model comparisons throughout the manuscript.

c) first analysis (p. 22, ls 24 and following). Given its relevance for the hypotheses, why not including the interaction term for the random slope since the beginning of the analysis?
AR: We have re-written this section. We have now motivated and specified the random effects structure at the beginning of the analysis (page 28), and used that throughout, changing only the fixed effects to test our hypotheses.

d) Tables reporting statistical results, please, specify in the legend what the numbers are (I assume the beta values and the standard errors (in parenthesis), but this should be spelt out).
AR: Values are specified in the revised results tables.

e) I think the readers may appreciate few lines commenting the baseline measure (i.e., Dutta's (2021) data). Moreover, the authors should clarify whether and how these baseline data were compared to the results of their study.
AR: We agree that this information could have been clearer in the original manuscript. We have now added a separate subsection in the introduction (page 16) discussing how we used the baseline Bengali measures, and more detailed speaker information for that data (c.f. response to comment 3 of reviewer 1).

6. I have found the discussion a bit unbalanced, as approximately half of it is dedicated to a secondary question (i.e., the impact of the paradigm on the transfer effects). I suggest the authors to re-balance the space, expanding the discussion of their main research question – they might further develop the theoretical implications of their findings, also offering some explanation of the results they consider unexpected.
AR: Thanks for this feedback; we have re-written the discussion section to include the suggested content (c.f. response to comment 3 of reviewer 1).

Minor

- p. 3 ls 39-43, To improve readiness, I would suggest the authors to consistently use only one term throughout the manuscript.
AH: We have modified the text to use the term "transfer" throughout, and specified this in the Introduction on page 3.

- in some places, the authors use capital letters to highlight some concepts, but I don't think they are really needed. Simple lower-case letters may be fine as well.
AR: We have edited these throughout the manuscript.

- Introduction, when discussing previous studies, the authors might clarify what acoustic measure each study has investigated (VOT, ...)
AR: We have added this information to citations throughout the introduction, and specified this on page 5.

- Acoustic analyses, third line, the authors report a total of 2333 target to be analyzed. How many of them were there for each target vowel?
AR: There were 1137 tokens of [ʌ] and 1196 tokens of [æ]; we have added this information into the text on page 24.

- p. 22, first line, please change “mediated” with “affected” or “moderated” or a similar term – mediated refers to mediation analysis, which is not the analysis run by the authors.
AR: We have made this edit.
